% Appendix C

\chapter{Licenced Software Usage}
\label{AppendixC}
\lhead{Appendix C. \emph{Licenced Software Usage}}

The Enigma application and associated utilities make use of several libraries and projects, all of which are used legally and in accordance with the matching software licence. The individual licence declarations are included within the files, or where appropriate. Below each library/project is listed, with their locations and other details.

\begin{enumerate}
	\item \textbf{Base64 Encoder/Decoder} \\
		\textbf{Author:} Robert Harder \\
		\textbf{Website:} http://www.iharder.net/current/java/base64/ \\
		\textbf{Licence:} None. Public domain release, full usage allowed. \\
		\textbf{Location:} \emph{com.cyanoryx.uni.common.Base64}
		
	\item \textbf{Cryptix Byte Utilities} \\
		\textbf{Author:} The Cryptix Foundation Ltd. \\
		\textbf{Website:} http://www.cryptix.org/ \\
		\textbf{Licence:} Cryptix General Licence -  http://www.cryptix.org/LICENSE.TXT \\
		\textbf{Location:} \emph{com.cyanoryx.uni.common.Bytes} \\
		\textbf{Notes:} Most methods only used partially.
		
	\item \textbf{Apache Xerces (XML parsing)} \\
		\textbf{Author:} Apache Software Foundation \\
		\textbf{Website:} http://xerces.apache.org/ \\
		\textbf{Licence:} Apache Licence, v2.0 - See licence at \emph{dep/APACHE-LICENSE-2.0.txt} \\
		\textbf{Location:} \emph{dep/xerces.jar}
		
	\item \textbf{Instant Messaging in Java} \\
		\textbf{Author:}  Iain Shigeoka (Manning) \\
		\textbf{Website:} http://xerces.apache.org/ \\
		\textbf{Licence:} The Shigeoka Software Licence - See licence at \emph{dep/SHIGEOKA-LICENCE.txt} \\
		\textbf{Location:} \emph{com.cyanoryx.uni.enigma.net.io.XercesReader.java}, plus occasional usage of code fragments and methods, which are noted within the code.
\end{enumerate}